---
title: "pathfindR Network Enrichment"
date: "4/20/2022"
output:
  html_notebook:
    highlight: default
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 2
    code: hide
  pdf_document:
    fig_caption: yes
    fig_crop: no
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---



```{r setup, include=FALSE}
library(pathfindR)
library(tidyverse)
library(dplyr)
```

```{r}
total_RNAseq <- readRDS(here::here("initial_data/total_RNAseq_all_annot_results.rds"))
network <- readRDS(here::here("initial_data/resistant_hybrid_network.rds"))
extended_network <- readRDS(here::here("initial_data/stringent_resistant.rds"))
```

# Start with the reduced triplicate network

# Create a dataframe of RNAseq data related to genes identified in network
```{r}
network_RNAseq <- semi_join(total_RNAseq, network, by = c("gene_name" = "mRNA"))
SK_R_RNAseq <- network_RNAseq %>% filter(grepl("SK", contrast))
SK_R_RNAseq <- select(SK_R_RNAseq, gene_name, log2FoldChange, pvalue)
extended_IPC_R_RNAseq <- network_RNAseq %>% filter(grepl("IPC", contrast))
extended_IPC_R_RNAseq <- select(extended_IPC_R_RNAseq, gene_name, log2FoldChange, pvalue)
```


# Run through various pathfindR PINs and databases

### Option to change the database (KEGG, Reactome, Biocarta, GO-BP, GO-All, GO-CC, GO-MF, cell_markers, mmu_KEGG, Custom (default: KEGG))
```{r}
output_ipc_Reactome <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "Reactome", output_dir = "results/triplicate_network/enrichment_results/IPCR_Reactome")
output_sk_Reactome <- run_pathfindR(SK_R_RNAseq, gene_sets = "Reactome", output_dir = "results/triplicate_network/enrichment_results/SKR_Reactome")
```



### Option to change the mapping PIN (Biogrid, STRING, GeneMania, IntAct, KEGG, mmu_STRING (default: biogrid))
```{r}
output_IPC_Reactome <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "Reactome", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/IPCR_Reactome")
output_SK_Reactome <- run_pathfindR(SK_R_RNAseq, gene_sets = "Reactome", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/SKR_Reactome")

output_IPC_GO_ALL <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "GO-All", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/IPCR_GO_ALL")
output_SK_GO_ALL <- run_pathfindR(SK_R_RNAseq, gene_sets = "GO-All", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/SKR_GO_ALL")

output_IPC_BioCarta <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/IPCR_BioCarta")
output_SK_BioCarta <- run_pathfindR(SK_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "STRING", output_dir = "results/triplicate_network/enrichment_results/SKR_BioCarta")

output_IPC_BioCarta_biogrid <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "Biogrid", output_dir = "results/triplicate_network/enrichment_results/IPCR_BioCarta_biogrid")
output_SK_BioCarta_biogrid <- run_pathfindR(SK_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "Biogrid", output_dir = "results/triplicate_network/enrichment_results/SKR_BioCarta_biogrid")
```




# Repeat with the extended network

# Create a dataframe of RNAseq data related to genes identified in network
```{r}
extended_network_RNAseq <- semi_join(total_RNAseq,extended_network, by = c("gene_name" = "mRNA"))
extended_SK_R_RNAseq <- extended_network_RNAseq %>% filter(grepl("SK", contrast))
extended_SK_R_RNAseq <- select(extended_SK_R_RNAseq, gene_name, log2FoldChange, pvalue)
extended_IPC_R_RNAseq <- extended_network_RNAseq %>% filter(grepl("IPC", contrast))
extended_IPC_R_RNAseq <- select(extended_IPC_R_RNAseq, gene_name, log2FoldChange, pvalue)
```


# Run through various pathfindR PINs and databases

### Option to change the database (KEGG, Reactome, Biocarta, GO-BP, GO-All, GO-CC, GO-MF, cell_markers, mmu_KEGG, Custom (default: KEGG))
```{r}
output_ipc_KEGG <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "KEGG", output_dir = "results/extended_network/enrichment_results/IPCR_KEGG")
output_sk_KEGG <- run_pathfindR(extended_SK_R_RNAseq, gene_sets = "KEGG", output_dir = "results/extended_network/enrichment_results/SKR_KEGG")
```



### Option to change the mapping PIN (Biogrid, STRING, GeneMania, IntAct, KEGG, mmu_STRING (default: biogrid))
```{r}
output_IPC_Reactome <- run_pathfindR(extended_extended_IPC_R_RNAseq, gene_sets = "Reactome", pin_name_path = "KEGG", output_dir = "results/extended_network/enrichment_results/IPCR_Reactome")
output_SK_Reactome <- run_pathfindR(extended_SK_R_RNAseq, gene_sets = "Reactome", pin_name_path = "KEGG", output_dir = "results/extended_network/enrichment_results/SKR_Reactome")

output_IPC_BioCarta <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "KEGG", output_dir = "results/extended_network/enrichment_results/IPCR_BioCarta")
output_SK_BioCarta <- run_pathfindR(extended_SK_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "KEGG", output_dir = "results/extended_network/enrichment_results/SKR_BioCarta")

output_IPC_GO_ALL <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "GO-All", pin_name_path = "STRING", output_dir = "results/extended_network/enrichment_results/IPCR_GO_ALL")
output_SK_GO_ALL <- run_pathfindR(extended_SK_R_RNAseq, gene_sets = "GO-All", pin_name_path = "STRING", output_dir = "results/extended_network/enrichment_results/SKR_GO_ALL")

output_IPC_BioCarta_biogrid <- run_pathfindR(extended_IPC_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "Biogrid", output_dir = "results/extended_network/enrichment_results/IPCR_BioCarta_biogrid")
output_SK_BioCarta_biogrid <- run_pathfindR(extended_SK_R_RNAseq, gene_sets = "BioCarta", pin_name_path = "Biogrid", output_dir = "results/extended_network/enrichment_results/SKR_BioCarta_biogrid")
```
