---
title: "total_network"
author: "Vincent Gureghian, Anthoula Gaigneaux, Hailee Herbst"
date: "4/4/2022"
output:
  html_notebook:
    highlight: default
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 2
    code: hide
  pdf_document:
    fig_caption: yes
    fig_crop: no
    number_sections: yes
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---


```{r setup}
library(tidyverse)
library(dplyr)
library(UpSetR)
library(ggupset)
library(igraph)
library(visNetwork)
library(linkcomm)
library(RCy3)

```

# Load data
```{r}
totalRNAseq <- readRDS(here::here("initial_data/total_RNAseq_all_annot_results.rds"))
hyb_sensitive <- readRDS(here::here("initial_data/hyb_sensitive.rds"))
hyb_resistant <- readRDS(here::here("initial_data/hyb_resistant.rds"))
smallRNAseq <- readRDS(here::here("initial_data/DEGs_smallRNAseq.rds"))

```

# Filter, separate, and compare totalRNAseq data
```{r pressure, echo=FALSE}

IPC_protein_coding <- totalRNAseq %>% 
  filter(grepl("protein_coding",biotype)) %>%
  filter(grepl("IPC298", contrast))

SKMel_protein_coding <- totalRNAseq %>% 
  filter(grepl("protein_coding",biotype)) %>%
  filter(grepl("SKMel30", contrast))

SK_IPC_comp <- semi_join(IPC_protein_coding, SKMel_protein_coding, by = "gene_name")
length(unique(SK_IPC_comp$gene_name))

```

# Filter, separate, and compare qCLASH data
``` {r}

IPC_hyb <- hyb_sensitive %>% filter(grepl("IPC298", cell_line))
SKMel_hyb <- hyb_sensitive %>% filter(grepl("SKMel30", cell_line))
IPC_R_hyb <- hyb_resistant %>% filter(grepl("IPC", Sample))
SKMel_R_hyb <- hyb_resistant %>% filter(grepl("SK30", Sample))


### Filter between dataframes based on hybrid presence ###

# compare SK to IPC
sensitive_comp <- semi_join(IPC_hyb, SKMel_hyb, by = "hybrids")
sensitive_comp <- distinct(sensitive_comp, hybrids, .keep_all = TRUE)
length(unique(sensitive_comp$hybrids))

# compare SK_R to IPC_R
resistant_comp <- semi_join(IPC_R_hyb, SKMel_R_hyb, by = "hybrids")
resistant_comp <- distinct(resistant_comp, hybrids, .keep_all = TRUE)
length(unique(resistant_comp$hybrids))
``` 


# Bridge RNAseq and qCLASH data
```{r}
dataset_resistant_comp <- semi_join(resistant_comp, SK_IPC_comp, by = c("mRNA" = "gene_name"))
dataset_sensitive_comp <- semi_join(sensitive_comp, SK_IPC_comp,by = c("mRNA" = "gene_name"))


dataset_resistant_differences <- anti_join(dataset_resistant_comp, dataset_sensitive_comp, by = "hybrids")
dataset_resistant_differences <- distinct(dataset_resistant_differences, hybrids, .keep_all = TRUE)
length(unique(dataset_resistant_differences$hybrids))
#saveRDS(dataset_resistant_differences, "resistant_differences.rds")

dataset_sensitive_differences <- anti_join(dataset_sensitive_comp, dataset_resistant_comp, by = "hybrids")
dataset_sensitive_differences <- distinct(dataset_sensitive_differences, hybrids, .keep_all = TRUE)
length(unique(dataset_sensitive_differences$hybrids))
#saveRDS(dataset_sensitive_differences, "sensitive_differences.rds")


```


# Build a network of miRNA-mRNA interactions for resistant cell lines 
``` {r}
qCLASH_treated <- readRDS(here::here("initial_data/hyb_resistant.rds"))
resistant_differences <-readRDS(here::here("initial_data/resistant_differences.rds"))
resistant_network <- semi_join(qCLASH_treated, resistant_differences, by = "hybrids")
```

```{r, eval=F}
#distribution among replicates

resistant_network %>% 
  group_by(Sample) %>% 
  count() %>% 
  ggplot(aes(x=Sample, y=n)) +
  geom_bar(stat="identity")

```



```{r, intersections of hybrids between resistant cell lines}
listInput <- list(IPC298_R = resistant_network %>% filter(grepl("IPC", Sample)) %>%  dplyr::select(hybrids) %>%  as_vector(),
                  SKMel30_R = resistant_network %>% filter(grepl("SK30",Sample)) %>% dplyr::select(hybrids) %>% as_vector())

upset(fromList(listInput), order.by = "freq")

```


```{r, intersections of hybrids between resistant technical replicates}

resistant_network %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(Sample=list(Sample)) %>% 
  ggplot(aes(x=Sample)) +
    geom_bar() +
    scale_x_upset(n_intersections = 36)
    #facet_wrap(~cell_line)

```



# Resistant Network construction

```{r, extract edges and build resistant network}

miRNA_gene_interaction <- resistant_network %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==6) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix()

res_network <- graph_from_edgelist(miRNA_gene_interaction, directed = FALSE)

visIgraph(res_network)


```

```{r, IPC_R network}
interactions_IPC298_R <- resistant_network %>% 
  filter(grepl("IPC", Sample)) %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix() 


IPC_resistant_network <- graph_from_edgelist (
  interactions_IPC298_R,
  directed = FALSE)

visIgraph(IPC_resistant_network)

```



```{r, SK_R network}

interactions_SKMel30_R <- resistant_network %>%
  filter(grepl("SK30", Sample)) %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix()


SK_resistant_network <- graph_from_edgelist (
  interactions_SKMel30_R,
  directed = TRUE)

visIgraph(SK_resistant_network)

```


# Build a network of miRNA-mRNA interactions for sensitive cell lines
```{r}
qCLASH_untreated <- readRDS(here::here("initial_data/hyb_sensitive.rds"))
sensitive_differences <-readRDS(here::here("initial_data/sensitive_differences.rds"))
sensitive_network <- semi_join(qCLASH_untreated, sensitive_differences, by = c("hybrids"))
```

```{r, eval=F}

#distribution among replicates
sensitive_network %>% 
  group_by(Sample) %>% 
  count() %>% 
  ggplot(aes(x=Sample, y=n)) +
  geom_bar(stat="identity")

```



```{r, intersections of hybrids between cell lines}
listInput <- list(IPC298 = sensitive_network %>% filter(cell_line=="IPC298") %>%  dplyr::select(hybrids) %>%  as_vector(),
                  SKMel30 = sensitive_network %>% filter(cell_line=="SKMel30") %>% dplyr::select(hybrids) %>% as_vector())

upset(fromList(listInput), order.by = "freq")

```


```{r, intersections of hybrids between technical replicates}

sensitive_network %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(Sample=list(Sample)) %>% 
  ggplot(aes(x=Sample)) +
    geom_bar() +
    scale_x_upset(n_intersections = 36)
    #facet_wrap(~cell_line)

```



# Sensitive Network construction

```{r, extract edges and build network}

miRNA_gene_interaction <- sensitive_network %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==6) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix()

sen_network <- graph_from_edgelist(miRNA_gene_interaction, directed = TRUE)

visIgraph(sen_network)

```

```{r, IPC network}

sensitive_network %>% 
  filter(cell_line == "IPC298") %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix() -> interactions_IPC298


IPC_sensitive_network <- graph_from_edgelist (
  interactions_IPC298,
  directed = TRUE)

visIgraph(IPC_sensitive_network)

```


```{r, SK network}

sensitive_network %>% 
  filter(cell_line == "SKMel30") %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix() -> interactions_SKMel30


SK_sensitive_network <- graph_from_edgelist (
  interactions_SKMel30,
  directed = TRUE)

visIgraph(SK_sensitive_network)

```


# Overlap IPC and SK data for sensitive and resistant networks to get lists of hybrids in each network
``` {r}
interactions_IPC298 <- as.data.frame.matrix(interactions_IPC298)
interactions_IPC298$hybrids <- paste(interactions_IPC298$miRNA, interactions_IPC298$mRNA, sep= "_")
interactions_SKMel30 <- as.data.frame.matrix(interactions_SKMel30)
interactions_SKMel30$hybrids <- paste(interactions_SKMel30$miRNA, interactions_SKMel30$mRNA, sep= "_")
sensitive_overlap <- semi_join(interactions_IPC298, interactions_SKMel30, by = "hybrids")
#saveRDS(sensitive_overlap, "sensitive_hybrid_network.rds")

interactions_IPC298_R <- as.data.frame.matrix(interactions_IPC298_R)
interactions_IPC298_R$hybrids <- paste(interactions_IPC298_R$miRNA, interactions_IPC298_R$mRNA, sep= "_")
interactions_SKMel30_R <- as.data.frame.matrix(interactions_SKMel30_R)
interactions_SKMel30_R$hybrids <- paste(interactions_SKMel30_R$miRNA, interactions_SKMel30_R$mRNA, sep= "_")
resistant_overlap <- semi_join(interactions_IPC298_R, interactions_SKMel30_R, by = "hybrids")
#saveRDS(resistant_overlap, "resistant_hybrid_network.rds")

```


# Conversion to Cytoscape Network
### Same network for IPC/SK - loaded into cytoscape individually to fit RNAseq data to the network
``` {r}
createNetworkFromIgraph(
 res_network,
  title = "SKMel30 Resistant Network",
  collection = "My Igraph Network Collection"
)

createNetworkFromIgraph(
  sen_network,
  title = "SKMel30 Sensitive Network",
  collection = "My Igraph Network Collection"
)

createNetworkFromIgraph(
 resistant_overlap,
  title = "IPC298 Resistant Network",
  collection = "My Igraph Network Collection"
)

createNetworkFromIgraph(
  sen_network,
  title = "IPC298 Sensitive Network",
  collection = "My Igraph Network Collection"
)
```



# Contrast Networks, unique for IPC_R and SK_R

# Anti-join dataframes to find unique hybrids of each resistant cell line
```{r} 
# hybrids in SK_R only, anti_join to the other 3 dataframes one by one # 
SK_R1 <- anti_join(SKMel_R_hyb, SKMel_hyb, by = "hybrids", .keep_all = TRUE)
SK_R2 <- anti_join(SK_R1, IPC_R_hyb, by = "hybrids", .keep_all = TRUE)
SK_R3 <- anti_join(SK_R2, IPC_hyb, by = "hybrids", .keep_all = TRUE)
SK_R_distinct <- distinct(SK_R3, hybrids, .keep_all = TRUE)
length(unique(SK_R_distinct$hybrids))

# hybrids in IPC_R only, anti_join to the other 3 dataframes one by one # 
IPC_R1 <- anti_join(IPC_R_hyb, IPC_hyb, by = "hybrids", .keep_all = TRUE)
IPC_R2 <- anti_join(IPC_R1, SKMel_R_hyb, by = "hybrids", .keep_all = TRUE)
IPC_R3 <- anti_join(IPC_R2, SKMel_hyb, by = "hybrids", .keep_all = TRUE)
IPC_R_distinct <- distinct(IPC_R3, hybrids, .keep_all = TRUE)
length(unique(IPC_R_distinct$hybrids))
```


# IPC_R unique network

# Data exploration
```{r, eval=F}
#distribution among replicates
IPC_R_network <- IPC_R3

IPC_R_network %>% 
  group_by(Sample) %>% 
  count() %>% 
  ggplot(aes(x=Sample, y=n)) +
  geom_bar(stat="identity")
```


```{r, intersections of hybrids between technical replicates for IPC_R}
IPC_R_network %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(Sample=list(Sample)) %>% 
  ggplot(aes(x=Sample)) +
    geom_bar() +
    scale_x_upset(n_intersections = 36)
    #facet_wrap(~cell_line)

```



# Network construction
```{r, IPC unique network}
interactions_IPC_R <- IPC_R_network %>% 
  filter(grepl("IPC", Sample)) %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix() 


IPC_R_unique_network <- graph_from_edgelist (
  interactions_IPC_R,
  directed = TRUE)

visIgraph(IPC_R_unique_network)
```


# SK_R unique network

# Data exploration
```{r, eval=F}
#distribution among replicates
SK_R_network <- SK_R3

SK_R_network %>% 
  group_by(Sample) %>% 
  count() %>% 
  ggplot(aes(x=Sample, y=n)) +
  geom_bar(stat="identity")
```


```{r, intersections of hybrids between technical replicates in SK_R unique}
SK_R_network %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(Sample=list(Sample)) %>% 
  ggplot(aes(x=Sample)) +
    geom_bar() +
    scale_x_upset(n_intersections = 36)
    #facet_wrap(~cell_line)
```


# Network construction
```{r, SK_R unique network}
interactions_SK_R <- SK_R_network %>% 
  filter(grepl("SK", Sample)) %>% 
  group_by(hybrids) %>% 
  summarise(samples = list(Sample %>% unique()), miRNA, mRNA ) %>%
  mutate(total = map_int( samples, length)) %>%
  distinct() %>% 
  filter(total==3) %>% 
  ungroup(hybrids) %>% 
  dplyr::select(miRNA, mRNA) %>% 
  as.matrix() 


SK_R_unique_network <- graph_from_edgelist (
  interactions_SK_R,
  directed = TRUE)

visIgraph(SK_R_unique_network)
```



```{r}
interactions_IPC_R <- as.data.frame.matrix(interactions_IPC_R)
interactions_IPC_R$hybrids <- paste(interactions_IPC_R$miRNA, interactions_IPC_R$mRNA, sep= "_")
#saveRDS(interactions_IPC_R, "IPC_R_hybrid_interactions.rds")

interactions_SK_R <- as.data.frame.matrix(interactions_SK_R)
interactions_SK_R$hybrids <- paste(interactions_SK_R$miRNA, interactions_SK_R$mRNA, sep= "_")
#saveRDS(interactions_SK_R, "SK_R_hybrid_interactions.rds")
```



# Conversion to Cytoscape Network
``` {r}
createNetworkFromIgraph(
 interactions_IPC_R,
  title = "Unique IPC298 Resistant Network",
  collection = "My Igraph Network Collection"
)

createNetworkFromIgraph(
  interactions_SK_R,
  title = "Unique SKMel30 Resistant Network",
  collection = "My Igraph Network Collection"
)


```


