---
title: "PCA of Stringent Threshold Network"
author: "Hailee Herbst"
date: "08/04/2022"
output:
  html_notebook:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(eulerr)
library(UpSetR)
library(ggupset)
library(DESeq2)
library(igraph)
library(visNetwork)
library(dplyr)
library(flextable)
```

__Formatting for paper__

```{r}
hyb_sensitive <- readRDS(here::here("extended_network_exploration/data/hyb_sensitive.rds"))
hyb_resistant <- readRDS(here::here("extended_network_exploration/data/hyb_resistant.rds"))
qCLASH <- bind_rows(hyb_sensitive, hyb_resistant)

thresh_qCLASH <- readRDS(here::here("extended_network_exploration/data/threshold_qCLASH.rds"))
MJ_qCLASH <- readRDS(here::here("extended_network_exploration/data/threshold_qCLASH_MJ.rds"))

```

# Tables

```{r}

tb <- thresh_qCLASH %>% 
  dplyr::select(cell_line, Sample,hybrids) %>%
  group_by(cell_line, Sample) %>% 
  mutate(total_hybrids=n()) %>%
  group_by(cell_line, Sample, hybrids) %>% 
  unique() %>%   
  group_by(cell_line, Sample) %>% 
  mutate(unique_hybrids=n()) %>% 
  dplyr::select(-hybrids) %>%
  unique() %>% 
  arrange(cell_line) %>% 
  dplyr::select("Cell line"=cell_line, Sample, "Total hybrids detected"= total_hybrids, "Unique hybrids detected"=unique_hybrids) %>% 
  flextable() 
save_as_image(tb, here::here("extended_network_exploration/data/qCLASH_table_hybrids_per_sample.png"))


```
# Upset Plots

### All Hybrids
```{r}
qCLASH %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(list_cell_line=list(cell_line)) %>% 
  ggplot(aes(x=list_cell_line)) +
    geom_bar() +
    scale_x_upset( n_intersections = 25)
```

### Threshold Hybrids
```{r}
thresh_qCLASH %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(list_cell_line=list(cell_line)) %>% 
  ggplot(aes(x=list_cell_line)) +
    geom_bar() +
    scale_x_upset( n_intersections = 25)
```

### Threshold Hybrids, MelJuso Data Included
```{r}
MJ_qCLASH %>% 
  dplyr::select(cell_line, Sample, hybrids) %>% 
  distinct(cell_line, Sample, hybrids, .keep_all=TRUE) %>%
  group_by(hybrids) %>% 
  summarize(list_cell_line=list(cell_line)) %>% 
  ggplot(aes(x=list_cell_line)) +
    geom_bar() +
    scale_x_upset( n_intersections = 25)
```

# Euler diagram for hybrids in biological replicates


### All Hybrids 
```{r}
list_hybrid <- function(.data=data, .replicate=replicate) {
  .data %>% filter(Sample==.replicate) %>% pull(hybrids) %>%  unique()
}

all_hybrids <- list(
           IPC1 = list_hybrid(qCLASH, "IPC1"),
           IPC2 = list_hybrid(qCLASH, "IPC2"),
           IPC3 = list_hybrid(qCLASH, "IPC3"),
           MJ1 = list_hybrid(qCLASH, "MJ1"),
           MJ2 = list_hybrid(qCLASH, "MJ2"),
           MJ3 = list_hybrid(qCLASH, "MJ3"),
           SK30_1 = list_hybrid(qCLASH, "SK30_1"),
           SK30_2 = list_hybrid(qCLASH, "SK30_2"),
           SK30_3 = list_hybrid(qCLASH, "SK30_3"),
           IPC_R1 = list_hybrid(qCLASH, "IPC_R1"),
           IPC_R2 = list_hybrid(qCLASH, "IPC_R2"),
           IPC_R3 = list_hybrid(qCLASH, "IPC_R3"),
           SK30_R1 =list_hybrid(qCLASH, "SK30_R1"),
           SK30_R2 = list_hybrid(qCLASH, "SK30_R2"),
           SK30_R3 = list_hybrid(qCLASH, "SK30_R3")
                                         )
ggarrange(plot(euler(all_hybrids[1:3], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(euler(all_hybrids[4:6], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(euler(all_hybrids[7:9], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(euler(all_hybrids[10:12], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(euler(all_hybrids[13:15], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5))
          )

ggarrange(plot(venn(all_hybrids[1:3]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[7:9]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[10:12]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[13:15]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5))
          )


```

### Threshold Hybrids

```{r}
list_hybrid <- function(.data=data, .replicate=replicate) {
  .data %>% filter(Sample==.replicate) %>% pull(hybrids) %>%  unique()
}

all_hybrids <- list(
           IPC1 = list_hybrid(thresh_qCLASH, "IPC1"),
           IPC2 = list_hybrid(thresh_qCLASH, "IPC2"),
           IPC3 = list_hybrid(thresh_qCLASH, "IPC3"),
           SK30_1 = list_hybrid(thresh_qCLASH, "SK30_1"),
           SK30_2 = list_hybrid(thresh_qCLASH, "SK30_2"),
           SK30_3 = list_hybrid(thresh_qCLASH, "SK30_3"),
           IPC_R1 = list_hybrid(thresh_qCLASH, "IPC_R1"),
           IPC_R2 = list_hybrid(thresh_qCLASH, "IPC_R2"),
           IPC_R3 = list_hybrid(thresh_qCLASH, "IPC_R3"),
           SK30_R1 =list_hybrid(thresh_qCLASH, "SK30_R1"),
           SK30_R2 = list_hybrid(thresh_qCLASH, "SK30_R2"),
           SK30_R3 = list_hybrid(thresh_qCLASH, "SK30_R3")
                                         )
ggarrange(plot(euler(all_hybrids[1:3], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[4:6], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[7:9], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[10:12], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )

ggarrange(plot(venn(all_hybrids[1:3]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[4:6]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[7:9]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[10:12]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )


```

### Threshold Hybrids, MelJuso Data Included

```{r}
list_hybrid <- function(.data=data, .replicate=replicate) {
  .data %>% filter(Sample==.replicate) %>% pull(hybrids) %>%  unique()
}

all_hybrids <- list(
           IPC1 = list_hybrid(MJ_qCLASH, "IPC1"),
           IPC2 = list_hybrid(MJ_qCLASH, "IPC2"),
           IPC3 = list_hybrid(MJ_qCLASH, "IPC3"),
           MJ1 = list_hybrid(MJ_qCLASH, "MJ1"),
           MJ2 = list_hybrid(MJ_qCLASH, "MJ2"),
           MJ3 = list_hybrid(MJ_qCLASH, "MJ3"),
           SK30_1 = list_hybrid(MJ_qCLASH, "SK30_1"),
           SK30_2 = list_hybrid(MJ_qCLASH, "SK30_2"),
           SK30_3 = list_hybrid(MJ_qCLASH, "SK30_3"),
           IPC_R1 = list_hybrid(MJ_qCLASH, "IPC_R1"),
           IPC_R2 = list_hybrid(MJ_qCLASH, "IPC_R2"),
           IPC_R3 = list_hybrid(MJ_qCLASH, "IPC_R3"),
           SK30_R1 =list_hybrid(MJ_qCLASH, "SK30_R1"),
           SK30_R2 = list_hybrid(MJ_qCLASH, "SK30_R2"),
           SK30_R3 = list_hybrid(MJ_qCLASH, "SK30_R3")
                                         )
ggarrange(plot(euler(all_hybrids[1:3], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[4:6], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[7:9], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[10:12], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(euler(all_hybrids[13:15], shape = "ellipse"), quantities = TRUE, fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )

ggarrange(plot(venn(all_hybrids[1:3]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[4:6]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[7:9]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[10:12]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[13:15]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )


```

#IPC
```{r}
list_hybrid <- function(.data=data, .replicate=replicate) {
  .data %>% filter(Sample==.replicate) %>% pull(hybrids) %>%  unique()
}

all_hybrids <- list(
           IPC1 = list_hybrid(qCLASH, "IPC1"),
           IPC2 = list_hybrid(qCLASH, "IPC2"),
           IPC3 = list_hybrid(qCLASH, "IPC3"),
           IPC_R1 = list_hybrid(qCLASH, "IPC_R1"),
           IPC_R2 = list_hybrid(qCLASH, "IPC_R2"),
           IPC_R3 = list_hybrid(qCLASH, "IPC_R3"),
           IPC1 = list_hybrid(thresh_qCLASH, "IPC1"),
           IPC2 = list_hybrid(thresh_qCLASH, "IPC2"),
           IPC3 = list_hybrid(thresh_qCLASH, "IPC3"),
           IPC_R1 = list_hybrid(thresh_qCLASH, "IPC_R1"),
           IPC_R2 = list_hybrid(thresh_qCLASH, "IPC_R2"),
           IPC_R3 = list_hybrid(thresh_qCLASH, "IPC_R3"))

ggarrange(plot(euler(all_hybrids[1:3], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), legend = NULL, alpha = 0.5)),
          plot(euler(all_hybrids[4:6], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5), legend = list(labels = c("IPC298 Replicate1", "IPC298 Replicate2", "IPC298 Replicate3"))),
          plot(euler(all_hybrids[7:9], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "palegreen4", "palegreen"), legend = NULL, alpha = 0.5)),
          plot(euler(all_hybrids[10:12], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5), legend = list(labels = c("IPC298_R Replicate1", "IPC298_R Replicate2", "IPC298_R Replicate3")))
          )

ggarrange(plot(venn(all_hybrids[1:3]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[4:6]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[7:9]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[10:12]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )

```

# SK
```{r}
list_hybrid <- function(.data=data, .replicate=replicate) {
  .data %>% filter(Sample==.replicate) %>% pull(hybrids) %>%  unique()
}

all_hybrids <- list(
           SK1 = list_hybrid(qCLASH, "SK30_1"),
           SK2 = list_hybrid(qCLASH, "SK30_2"),
           SK3 = list_hybrid(qCLASH, "SK30_3"),
           SK_R1 = list_hybrid(qCLASH, "SK30_R1"),
           SK_R2 = list_hybrid(qCLASH, "SK30_R2"),
           SK_R3 = list_hybrid(qCLASH, "SK30_R3"),
           SK1 = list_hybrid(thresh_qCLASH, "SK30_1"),
           SK2 = list_hybrid(thresh_qCLASH, "SK30_2"),
           SK3 = list_hybrid(thresh_qCLASH, "SK30_3"),
           SK_R1 = list_hybrid(thresh_qCLASH, "SK30_R1"),
           SK_R2 = list_hybrid(thresh_qCLASH, "SK30_R2"),
           SK_R3 = list_hybrid(thresh_qCLASH, "SK30_R3"))

plot <-ggarrange(plot(euler(all_hybrids[1:3], shape = "ellipse"), quantities = list(type = "percent"),  fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5), legend = NULL),
          plot(euler(all_hybrids[4:6], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5), legend = list(labels = c("SKMel30 Replicate1", "SKMel30 Replicate2", "SKMel30 Replicate3"))),
          plot(euler(all_hybrids[7:9], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5), legend = NULL),
          plot(euler(all_hybrids[10:12], shape = "ellipse"), quantities = list(type = "percent"), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5), legend = list(labels = c("SKMel30_R Replicate1", "SKMel30_R Replicate2", "SKMel30_R Replicate3")))
          ) 

ggarrange(plot(venn(all_hybrids[1:3]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[4:6]), fills = list(fill = c("snow", "steelblue3", "lightsteelblue"), alpha = 0.5)),
          plot(venn(all_hybrids[7:9]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5)),
          plot(venn(all_hybrids[10:12]), fills = list(fill = c("snow", "palegreen4", "palegreen"), alpha = 0.5))
          )

plot

```


# PCA 

### PCA for Hybrids, All Data Included

```{r, message=FALSE, warning=FALSE}
counts1 <- readRDS(here::here("extended_network_exploration/data/hybrid_as_count_table.RDS"))
rownames(counts1) <- counts1$hybrids
counts1 <- counts1 %>% select(-hybrids)
                            
coldata1 <- tibble(condition = factor(rep(c("untreated","untreated", "untreated", "treated", "treated"), each = 3, len = 15), 
                                 levels = c("untreated", "treated")),
               cell_line = factor(rep(c("IPC298","SKMel30", "MelJuso", "IPC298", "SKMel30"),each = 3, len = 15)),
               samples = factor(paste(rep(c("IPC", "SK", "MJ","IPC", "SK"),each = 3,len = 15),
                                    c(rep(c("1", "2", "3"),len=9), rep(c("R1", "R2", "R3"), len=6)), sep = "_")),
            
               )

dds1 <- DESeqDataSetFromMatrix(countData = counts1,
                              colData   = coldata1,
                              design = ~condition)

dds1 <- estimateSizeFactors(dds1)
```

```{r}
make_PCA_ntop500 <- function(.dds1) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds1, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 500, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 500 - Original qCLASH Data PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop500(.dds1=dds1)
```

```{r}
make_PCA_ntop1000 <- function(.dds1) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds1, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 1000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 1,000 - Original Data PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop1000(.dds1=dds1)
```

```{r}
make_PCA_ntop10000 <- function(.dds1) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds1, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 10000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 10,000 - Original Data PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop10000(.dds1=dds1)
```


### PCA for Hybrids Which Pass Threshold of Expression
```{r, message=FALSE, warning=FALSE}
counts2 <- readRDS(here::here("extended_network_exploration/data/count_table_threshold.RDS"))
rownames(counts2) <- counts2$hybrids
counts2 <- counts2 %>% select(-hybrids)
counts2[is.na(counts2)] <- 0                            
                            
coldata2 <- tibble(condition = factor(rep(c("untreated","untreated", "treated", "treated"), each = 3, len = 12), 
                                 levels = c("untreated", "treated")),
               cell_line = factor(rep(c("IPC298","SKMel30", "IPC298", "SKMel30"),each = 3, len = 12)),
               samples = factor(paste(rep(c("IPC", "SK", "IPC", "SK"),each = 3,len = 12),
                                    c(rep(c("1", "2", "3"),len=6), rep(c("R1", "R2", "R3"), len=6)), sep = "_")),
            
               )

dds2 <- DESeqDataSetFromMatrix(countData = counts2,
                              colData   = coldata2,
                              design = ~condition)

dds2 <- estimateSizeFactors(dds2)
```


```{r}
make_PCA_ntop500 <- function(.dds2) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds2, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 500, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 500 - Threshold Network Data PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop500(.dds2=dds2)
```


```{r}
make_PCA_ntop1000 <- function(.dds2) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds2, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 1000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 1,000 - Threshold PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop1000(.dds2=dds2)
```
```{r}
make_PCA_ntop10000 <- function(.dds2) {
  
vsd_NotBlind <- varianceStabilizingTransformation(.dds2, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop= 10000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 10,000 - Threshold PCA")+
    theme(plot.title = element_text(hjust = 0.5))
}

make_PCA_ntop10000(.dds2=dds2)
```

### PCA with MelJuso Data (sensitive only) for Hybrids Which Pass Threshold

```{r, message=FALSE, warning=FALSE}

counts3 <- readRDS(here::here("extended_network_exploration/data/count_table_threshold_MJ.RDS"))
rownames(counts3) <- counts3$hybrids
counts3 <- counts3 %>% select(-hybrids)
counts3[is.na(counts3)] <- 0                            
                            
coldata3 <- tibble(condition = factor(rep(c("untreated","untreated", "untreated", "treated", "treated"), each = 3, len = 15), 
                                 levels = c("untreated", "treated")),
               cell_line = factor(rep(c("IPC298","SKMel30", "MelJuso", "IPC298", "SKMel30"),each = 3, len = 15)),
               samples = factor(paste(rep(c("IPC", "SK", "MJ","IPC", "SK"),each = 3,len = 15),
                                    c(rep(c("1", "2", "3"),len=9), rep(c("R1", "R2", "R3"), len=6)), sep = "_")),
            
               )

 dds3 <- DESeqDataSetFromMatrix(countData = counts3,
                              colData   = coldata3,
                              design = ~condition)

dds3 <- estimateSizeFactors(dds3)
```



```{r}

make_PCA_ntop500 <- function(.dds3) {
  
vsd_NotBlind <- varianceStabilizingTransformation(dds3, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop = 500, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 500 - Threshold Network Data PCA")+
    theme(plot.title = element_text(hjust = 0.5))

}

make_PCA_ntop500(.dds3 = dds3)

```


```{r}

make_PCA_ntop1000 <- function(.dds3) {
  
vsd_NotBlind <- varianceStabilizingTransformation(dds3, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop = 1000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 1,000 - Threshold PCA with MelJuso")+
    theme(plot.title = element_text(hjust = 0.5))

}

make_PCA_ntop1000(.dds3 = dds3)

```
```{r}

make_PCA_ntop10000 <- function(.dds3) {
  
vsd_NotBlind <- varianceStabilizingTransformation(dds3, blind=FALSE)
pca_df <- plotPCA(vsd_NotBlind, ntop = 10000, intgroup = c("condition", "cell_line"), returnData = TRUE)
percentVar <- round(100 * attr(pca_df, "percentVar"))
pca_df %>%  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(shape = cell_line,  color = condition), size = 3, alpha = .8) +
  #ggrepel::geom_text_repel(aes(label = name), 
  #                         min.segment.length = .1, show.legend = FALSE) +
  ggforce::geom_mark_ellipse(aes(label = cell_line, group = cell_line)) +
  theme_bw(12) +
  #facet_wrap(~ cell, scales = "free") +
  labs(x = paste0("PC1: ", percentVar[1], "% variance"),
       y = paste0("PC2: ", percentVar[2], "% variance"),
       title = "Top 10,000 - Threshold PCA with MelJuso")+
    theme(plot.title = element_text(hjust = 0.5))

}

make_PCA_ntop10000(.dds3 = dds3)

```




